---
title: "House Price Predictor"
author: "Sajudra Gupta"
date: "`r Sys.Date()`"
output: html_document
---

## Calling of required libraries
```{r}
library(dplyr)
library(tidyr)
library(caret)
library(rpart)
library(base)
library(corrplot)
```

## Loading of required data
```{r}
house_prices <- read.csv("C:/Users/gupta/OneDrive/Desktop/ML Projects/House Price Predictor/house_prices.csv")
```

## Data Preprocessing
```{r}
# Information in table
head(house_prices)
dim(house_prices)

#Summary of table
summary(house_prices)
house_prices [house_prices == ""|house_prices == " "] <- NA #Converting empty cells into NAs
colSums(is.na(house_prices))

#Removal of variables having more than 50% NAs
pricing <- house_prices[, colSums(is.na(house_prices))/nrow(house_prices) <0.5 ]
pricing$Index<- NULL

#head(pricing)
head(pricing)
summary(pricing)
```

## Feautre Engineering

1. Converting numeric strings to numeric
```{r}
pricing$Balcony <-as.numeric(pricing$Balcony)
pricing$Bathroom <-as.numeric(pricing$Bathroom)
```

2. Converting lakh and crores into numerics values
```{r}
pricing$Amount.in.rupees. <- as.numeric(gsub("Lac | Cr", " ", pricing$Amount.in.rupees.))* # Detecting Lac and Cr in Amount column
  ifelse(grepl("Cr", pricing$Amount.in.rupees.), 100, 1)*100000 #Replacing the words with values with 100000 (100x for Cr, 1x for Lac)
```


3. Changing floor into fraction (calling it floor ratio)
```{r}
pricing$floor_ratio<- ifelse(grepl("Basement|upper|Lower", pricing$Floor, ignore.case=TRUE),-1, NA) #Assuming any level of basement: -1
pricing$Floor <- sub(".*?(\\d+)\\s*out\\s*of\\s*(\\d+).*", "\\1/\\2",pricing$Floor) #Bringing the floors in a fraction format 
split_floors <- strsplit(pricing$Floor, "/")
pricing$floor_ratio <- as.numeric(sapply(split_floors, `[`, 1)) / as.numeric(sapply(split_floors, `[`, 2)) #Converting and evaluating the fractions from string to numerics
pricing$floor_ratio[grepl("-1", pricing$Floor, ignore.case=TRUE)]<- -1 # Changing the data type from string to numeric
pricing$floor_ratio[grepl("Ground", pricing$Floor, ignore.case=TRUE)] <- 0 #Changing ground floors to zero
pricing$floor_ratio[pricing$floor_ratio > 1] <- 1 #Assigning roof tops and penthouses to 1
```

4. Extracting no. of BHKs
```{r}
pricing$bhk <- ifelse(grepl("Studio|Apartment", pricing$Title, ignore.case = TRUE),0.5,as.numeric(sub(".*?(\\d+)\\s*BHK.*", "\\1", pricing$Title))) #Setting studios for 0.5 BHK and others as per their given formula
```

5. Removing units from area
```{r}
pricing$Carpet.Area <- as.numeric(sub("(\\d+)\\s*sqft.*", "\\1", pricing$Carpet.Area)) #Removing the units and make the value numeric
```


##DATA CLEANING

1. Removing NAs with median for numeric and mode for character
```{r}
pricing <- pricing %>%
  mutate(across(where(is.numeric), #detects numeric variables
                ~replace_na(., median(., na.rm = TRUE)))) %>% #replaces numeric NAs with median
  mutate(across(where(is.character), ~{ #detects character variables
    mode_val <- names(sort(table(.), decreasing = TRUE))[1] #sorting names in decreasing order wrt freqency, and extracting the first name
    replace_na(., mode_val) #replaces NAs with mode
  }))
anyNA(pricing)
head(pricing)
```


2. Detecting outliers
```{r}
summary(pricing$floor_ratio)
summary(pricing$bhk)
summary(pricing$Balcony)
summary(pricing$Bathroom)
```


```{r}
#Function for outlier detection
oultier_detector <-function(data, column){
  if(is.data.frame(data)) {
    col_data <- data[[column]]
  } else {
    col_data <- data  
  }
  
  q1 <- quantile(col_data,0.25,na.rm = TRUE)
  q3 <- quantile(col_data,0.75,na.rm = TRUE)
  iqr<- q3 - q1
  upper <- q3+1.5*iqr
  lower <- q1-1.5*iqr
  
  n_outliers <- sum(col_data < lower | col_data > upper)
  return(n_outliers)
}
```

```{r}
#Function to cap outliers
oultier_capper <- function(data, column){
  if(is.data.frame(data)) {
    col_data <- data[[column]]
  } else {
    col_data <- data  
  }
  q1 <- quantile(col_data,0.25,na.rm = TRUE)
  q3 <- quantile(col_data,0.75,na.rm = TRUE)
  iqr<- q3 - q1
  upper <- q3+1.5*iqr
  lower <- q1-1.5*iqr
  
  capped_data <- col_data
  capped_data[col_data < lower] <- lower
  capped_data[col_data > upper] <- upper
  return(invisible(capped_data))
}
```


```{r}
oultier_detector(pricing$Price..in.rupees.)
```

```{r}
pricing$Price..in.rupees.<- oultier_capper(pricing$Price..in.rupees.)
summary(pricing$Price..in.rupees.)
```

3. Removing Title and Description
```{r}
pricing$Title <- NULL
pricing$Description <- NULL #Since Bhk is extracted, they are of no use anymore
pricing$Amount.in.rupees. <- NULL #Focusing on prices rather than indivdual amount 
pricing$Status <- NULL #Since all are "ready to move"
pricing$facing<- NULL
```


## EXPLORATORY DATA ANALYSIS
```{r}
hist(pricing$floor_ratio)
hist(pricing$bhk)
hist(pricing$Price..in.rupees.)
```

```{r}
#Log transform of right skewed data
pricing$log_bhk <- log(pricing$bhk)
pricing$log_Price..in.rupees. <- log(pricing$Price..in.rupees. + 1)
```

```{r}
hist(pricing$log_bhk )
hist(pricing$log_Price..in.rupees. )
```

```{r}
nums <- sapply(pricing, is.numeric)
cor_price <- cor(pricing[, nums], use = "complete.obs")

round(cor_price, 2)
corrplot(cor_price)
pricing$overlooking <- as.factor(pricing$overlooking)
```

We can observe that, the prices depend more on the no. of the available bathrooms and the balcony than the floor or the bhk of the house. 

## REGRESSION MODEL

1. Linear Regression
```{r}
model_lmb <- lm(log_Price..in.rupees. ~location + floor_ratio + Transaction + Furnishing + log_bhk + Bathroom + Balcony + Ownership+overlooking, data = pricing)
summary(model_lmb)
```

```{r}
AIC(model_lmb)
```

```{r}
split_data <- function(pricing) # For data splitting
{
  train_split = createDataPartition(pricing$Price..in.rupees., p = 0.8, list = FALSE)
  train_price = pricing[train_split,]
  test_price = pricing[-train_split,]
  return(list(train_price,test_price))
}
pricing_split <- split_data(pricing)
train_pricing <- pricing_split[[1]]
test_pricing <- pricing_split[[2]]
pricing$overlooking<-as.factor(pricing$overlooking)
pricing$Transaction <- as.factor(pricing$Transaction)
```

```{r}
model_btr <- lm(Price..in.rupees.~location + floor_ratio + Transaction + Furnishing + bhk + Bathroom + Balcony + Ownership+overlooking+Carpet.Area, data = train_pricing)
summary(model_btr)
model_ltbr<-lm(log_Price..in.rupees.~location + floor_ratio + Transaction + Furnishing + log_bhk + Bathroom + Balcony + Ownership+overlooking+Carpet.Area, data = train_pricing)
summary(model_ltbr)
```

```{r}
#Calculating MAE
pred_price_lr <- predict(model_btr, data = test_pricing)
mae_lr <- mean(abs(test_pricing$Price..in.rupees. - pred_price_lr))
mae_lr

#Calculating RMSE
rmse_lr<- RMSE(test_pricing$Price..in.rupees.,pred_price_lr)
rmse_lr

#Calculating AIC
AIC(model_btr)
```

2. Random forest model

```{r}
library(randomForest)

model_rf <- randomForest(Price..in.rupees.~location + floor_ratio + Transaction + Furnishing + bhk + Bathroom + Balcony + Ownership+overlooking+Carpet.Area, data = train_pricing,
                        ntree = 500,      # Number of trees
                        importance = TRUE, # Feature importance
                        nodesize = 500,
                        sampsize = 50000) # Min observations per leaf
```

```{r}
print(model_rf)
```


```{r}
# Predict & evaluate
pred_rf <- predict(model_rf, test_pricing)
rmse_rf <- sqrt(mean((test_pricing$Price..in.rupees. - pred_rf)^2))
mae_rf <- mean(abs(test_pricing$Price..in.rupees. - pred_rf))
mae_rf
rmse_rf
test_price <- test_pricing$Price..in.rupees.
r2_rftest<- 1-sum((test_price - pred_rf)^2) / sum((test_price - mean(test_actual))^2)
r2_rftest
```

```{r}
price_model <- model_rf
saveRDS(price_model, "House Price Predictior.rds")
```


